Methodology for Assessing the State of the Practice for Domain X

### 2020-10-15 StateOfPractice/Methodology/Methodology.tex
added another commonality analysis reference to Section 7 of Methodology
fixed type Section 7 of Methodology
updated methodology with meeting comments

\begin{enumerate}
\item What artifacts do current software packages produce? 
\item What tools are used by current software packages?
\item What principles, processes, and methodologies are used in the development of current software packages?
\item What are the pain points for developers working on SCS projects? What aspects of the existing processes, methodologies and tools do they consider could potentially be improved? What processes, methodologies and tools have they changed in order to reduce obstacles?
\item What are the attitudes and actions of research software developers toward:
\begin{enumerate}
\item sustainability?
\item usability?
\item modifiability?
\end{enumerate} 
\item For a given domain of research software, what current projects provide the best examples of high quality software and documentation?
\end{enumerate}

@@

\begin{enumerate}
\item Start with the state of practice research questions. (Section~\ref{ResearchQuestions}) \pmi{To be updated based on feedback}
\item Identify the domain. (Section~\ref{SecIdentifyDomain}) \pmi{In second review}
\item \emph{Domain Experts}: Create a top ten list of software packages in the domain. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Brief the Domain Experts on the overall objective, research proposal, research questions, measurement template, survey for short list projects, usability tests, performance benchmarks, maintainability experiments. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Identify broad list of candidate software packages in the domain. (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages list. (Section~\ref{SecInitialFilter}) \pmi{To be edited based on feedback}
\item \emph{Domain Experts}: Review domain software list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})\pmi{To be edited - add intro}
\item \emph{Domain Experts}: Vet domain analysis. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts}) 
\item Gather source code and documentation for each prospective software package.
\item Collect empirical measures. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecShallowMeasure}) \pmi{To be edited based on feedback}
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
\item Identify a short list of top software packages, typically four to six, for deeper exploration according to the AHP rankings of the measurements.
\item \emph{Domain Experts}: Vet AHP ranking and short list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item With short list:
\begin{enumerate}
\item Survey developers (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions to Developers.pdf}{Questions to Developers})
\item Usability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/User Experiments.pdf}{User Experiments})
\item Performance benchmarks\pmi{note: this is still in consideration}
\item Maintainability experiments\pmi{note: these experiments need to be completed - to be discussed at next meeting}
\end{enumerate}
\item Rank short list. (Section~\ref{SecRankShortList}) \pmi{To be completed}
\item Document answers for research questions.
\end{enumerate}

\wss{The domain expert is involved in multiple steps in the process.  How best
  to get their feedback?  The domain experts are busy and are unlikely to devote
  significant time to the project.  We need to quickly get to the point.  Maybe
  something around task based inspection?  Directed interview?}

@@

The domain analysis consists of a commonality analysis of the family of software packages. Its purpose is to show the relationships between these packages, and to facilitate an understanding of the informal specification and development of them. \cite{weiss1997defining} defines commonality analysis as an approach to defining a family by identifying commonalities, variabilities, and common terminology for the family. Commonalities are goals, theories, models, definitions and assumptions that are common between family members. Likewise, variabilities are goals, theories, models, definitions and assumptions that are different between family members. The final result of the domain analysis will be tables of commonalities, variabilities, and parameters of variation of a program family, which \cite{parnas1976design} defines as ``a set of programs whose common properties are so extensive that it is advantageous to study the common properties of the programs before analyzing individual members''. \cite{smith2008commonality} present a template for conducting a commonality analysis, which was referred to when conducting this work. \cite{weiss1998commonality} describes another commonality analysis technique for deciding the members of a program family which readers are encouraged to be familiar with. \cite{SmithAndChen2004} provide an example of a commonality analysis for mesh generating software. The steps to produce a commonality analysis are:

>>>>

\begin{enumerate}
\item What are the artifacts that current software packages consist of? 
\item What tools are used by current software packages?
\item What principles, processes, and methodologies are used in the development of current software packages?
\item What are the pain points for developers working on SCS projects? What aspects of the existing processes, methodologies and tools do they consider could potentially be improved? What processes, methodologies and tools have they changed in order to reduce obstacles?
\item What are the actions of research software developers toward:\pmi{to do: reverse engineer from questions to developers; might remove or completely alter points below}
\begin{enumerate}
\item sustainability?\pmi{meeting discussion did consider removing this}
\item usability?
\item modifiability?
\end{enumerate} 
\item How does software designated as high quality by this methodology compare with top rated software by the community? \pmi{review}
\end{enumerate}

@@

\begin{enumerate}
\item Start with the state of practice research questions. (Section~\ref{ResearchQuestions}) \pmi{To be updated based on feedback}
\item Identify the domain. (Section~\ref{SecIdentifyDomain}) \pmi{In second review}
\item \emph{Domain Experts}: Create a top ten list of software packages in the domain. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Brief the Domain Experts on the overall objective, research proposal, research questions, measurement template, survey for short list projects, usability tests, performance benchmarks, maintainability experiments. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Identify broad list of candidate software packages in the domain. (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages list. (Section~\ref{SecInitialFilter}) \pmi{To be edited based on feedback}
\item \emph{Domain Experts}: Review domain software list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})\pmi{To be edited - add intro}
\item \emph{Domain Experts}: Vet domain analysis. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts}) 
\item Gather source code and documentation for each prospective software package.
\item Collect empirical measures. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecShallowMeasure}) \pmi{To be edited based on feedback}
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
\item Identify a short list of top software packages, typically four to six, for deeper exploration according to the AHP rankings of the measurements.
\item \emph{Domain Experts}: Vet AHP ranking and short list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item With short list:
\begin{enumerate}
\item Survey developers (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions to Developers.pdf}{Questions to Developers})
\item Usability experiments (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/User Experiments.pdf}{User Experiments})
\item Performance benchmarks\pmi{note: this is still in consideration}
\item Maintainability experiments\pmi{note: these experiments need to be completed - to be discussed at next meeting}\pmi{meeting: start thinking about maintainability experiments; maybe change to modifiability; at least describe what we would do for such experiments - measure sensible changes, what does a small change mean exactly, document 'likely changes' - see Parnas; modifiability - we want to be explicit about what we want to modify; how to we interpret the results?; Add these experiments to user experiments file, in such case change file name, keep format similar to user experiments (procedure, no questionnaire, add preamble describing comments above), such experiments would get quantitative measures}\pmi{SS:For the maintainability experiment, I think we should write it up in the methodology, even if we don't do it. Discussing maintainability is needed to fit with the research questions (since maintainability is part of sustainability). We want sustainability to fit into the bigger context of what interests research software developers. As far as the maintainability experiment goes, we can just ask the domain experts to make a likely change and ask them to record their experience with making the change. It isn't quantifiable, but it would give us qualitative data.}
\end{enumerate}
\item Rank short list. (Section~\ref{SecRankShortList}) \pmi{To be completed}
\item Document answers for research questions.
\end{enumerate}

%\wss{The domain expert is involved in multiple steps in the process.  How best
%  to get their feedback?  The domain experts are busy and are unlikely to devote
%  significant time to the project.  We need to quickly get to the point.  Maybe
%  something around task based inspection?  Directed interview?}

@@

The domain analysis consists of a commonality analysis of the family of software packages. Its purpose is to show the relationships between these packages, and to facilitate an understanding of the informal specification and development of them. \cite{weiss1997defining} defines commonality analysis as an approach to defining a family by identifying commonalities, variabilities, and common terminology for the family. Commonalities are goals, theories, models, definitions and assumptions that are common between family members. Likewise, variabilities are goals, theories, models, definitions and assumptions that are different between family members. The final result of the domain analysis will be tables of commonalities, variabilities, and parameters of variation of a program family, which \cite{parnas1976design} defines as ``a set of programs whose common properties are so extensive that it is advantageous to study the common properties of the programs before analyzing individual members''. \cite{smith2008commonality} present a template for conducting a commonality analysis, which was referred to when conducting this work. \cite{weiss1998commonality} describes another commonality analysis technique for deciding the members of a program family which readers are encouraged to be familiar with. \cite{SmithAndChen2004} and \cite{SmithMcCutchanAndCarette2017} are examples of a commonality analysis for a family of mesh generating software and a family of material models, respectively. The steps to produce a commonality analysis are:
